Flexibility of interactions depends on habitat suitability

### 2019-07-20 paper/manuscript.Rmd
Say I added GBIF with network occurrences
Start drafting how niche suitability is calculated

We retrieved occurrences from the Global Biodiversity Information Facility (GBIF; https://www.gbif.org) using the R package `rgbif` `r as.character(packageVersion('taxize'))` [@chamberlain_r_2017; @chamberlain_rgbif_2019].
For each species, we downloaded occurrences for which the coordinates of the observation were available and that had no known geospatial issue in the GBIF database.
Roughly, we downloaded `r round(sum(n_occurrences$N)/1000000, digits = 1)` million occurrences for `r format(n_species_with_occurrences, big.mark = ",")` of the species of interest. 
The occurrences obtained contain observations of mixed quality, so we applied a series of simple filters to remove those of lower quality. 
\R{rationale-occurrence-filtering}Specifically, we removed all occurrences with (*i*) a coordinate uncertainity larger than 100km; (*ii*) those recorded prior to 1945 (ref? or why?); (*iii*) those in which the number of individuals recorded was either zero or more than 100 as such large counts are often erroneous (ref? or why?); and (*iv*) those occurrences in which the "basis of record" was not a human observation or a preserved specimen (ref? why?). 
We then used the R package `CoordinateCleaner` `r as.character(packageVersion('taxize'))` [@zizka_coordinatecleaner_2019] and land mass and country data from Natural Earth (https://www.naturalearthdata.com) with a 1:10,000,000 scale to further identify and remove problematic occurrences. 
We removed occurrences in which their coordinates (*v*) fell outside the borders of the country in which they were recorded; (*vi*) those arround a the country capital or the centroid of the country and province centroids; (*vii*) those arond a biodiversity institution; and (*viii*) those located within oceans. 
Thorugh this cleaning process, we removed with `r round(sum(n_dirty_occurrences$N)/1000000, digits = 1)` million occurrences distributed accross `r n_species_with_dirty_occurrences` species.

Next, we overlapped the occurrence and our environmental rasters such that each occurrence location is characterized by a vector with information of our 37 bioclimatic and topographic variables . 
If a species had more than one occurrence records within one of the cells of the bioclimatic raster we only included one of the occurrence records. 
We did this to avoid giving more weight to areas with a high number of occurrences, a common scenario in occurrrence records collected opportunistically as the ones we use here. 
Overall we removed `r scales::percent(percent_thinned)` of the occurrences which resulted in a total of `r round(sum(n_thinned_occurrences$N)/1000000, digits = 1)` million occurrences used in our niche analysis. 
A common issue of terrestrial bioclimatic datasets is that the boundaries of the cells with information do not exactly match the land mass boundaries. 
The result of this missmatch is that at least one of the environmental variables was not available for `r format(complete_climate_cells[1], big.mark = ",")` of the raster cells with occurrences (`r scales::percent(complete_climate_cells[1]/sum(complete_climate_cells), accuracy = 0.1)` of the total). 
As expected a vast majority of these cells were close to the shore. 
To address this issue, calculated average value of environmental variables within a `r config$climate_buffer/1000`km buffer of the center of the cell where the variable is missing, and used it to approximate the value of the variable in that cell. 
In this fashion we were able to fill environmental variables for `r scales::percent(percent_filled_round_one, accuracy = 0.1)` of the cells were they were missing. 
To fill the remaining `r format(complete_filled_cells_1[1], big.mark= ",")` cells, we repeated the aforementioned procedure but using a `r config$climate_buffer*2/1000`km buffer. 
We removed from further analyis occurrences located within the `r complete_filled_cells_2[1]` cells for which we were unable to fill environmental variables.

>>>>

We retrieved occurrences from the Global Biodiversity Information Facility (GBIF; https://www.gbif.org) using the R package `rgbif` `r as.character(packageVersion('taxize'))` [@chamberlain_r_2017; @chamberlain_rgbif_2019].
For each species, we downloaded occurrences for which the coordinates of the observation were available and that had no known geospatial issue in the GBIF database.
Roughly, we downloaded `r round(sum(n_occurrences$N)/1000000, digits = 1)` million occurrences for `r format(n_species_with_occurrences, big.mark = ",")` of the species of interest. 
The occurrences obtained contain observations of mixed quality, so we applied a series of simple filters to remove those of lower quality. 
\R{rationale-occurrence-filtering}Specifically, we removed all occurrences with (*i*) a coordinate uncertainity larger than 100km; (*ii*) those recorded prior to 1945 (ref? or why?); (*iii*) those in which the number of individuals recorded was either zero or more than 100 as such large counts are often erroneous (ref? or why?); and (*iv*) those occurrences in which the "basis of record" was not a human observation or a preserved specimen (ref? why?). 
We then used the R package `CoordinateCleaner` `r as.character(packageVersion('taxize'))` [@zizka_coordinatecleaner_2019] and land mass and country data from Natural Earth (https://www.naturalearthdata.com) with a 1:10,000,000 scale to further identify and remove problematic occurrences. 
We removed occurrences in which their coordinates (*v*) fell outside the borders of the country in which they were recorded; (*vi*) those arround a the country capital or the centroid of the country and province centroids; (*vii*) those arond a biodiversity institution; and (*viii*) those located within oceans. 
Thorugh this cleaning process, we removed with `r round(sum(n_dirty_occurrences$N)/1000000, digits = 1)` million occurrences distributed accross `r n_species_with_dirty_occurrences` species. 
We added the occurrences from our plant-pollinator networks to the cleaned occurrences retrieved from GBIF.

Next, we overlapped the occurrence and our environmental rasters such that each occurrence location is characterized by a vector with information of our 37 bioclimatic and topographic variables . 
If a species had more than one occurrence records within one of the cells of the bioclimatic raster we only included one of the occurrence records. 
We did this to avoid giving more weight to areas with a high number of occurrences, a common scenario in occurrrence records collected opportunistically as the ones we use here. 
Overall we removed `r scales::percent(percent_thinned)` of the occurrences which resulted in a total of `r round(sum(n_thinned_occurrences$N)/1000000, digits = 1)` million occurrences used in our niche analysis. 
A common issue of terrestrial bioclimatic datasets is that the boundaries of the cells with information do not exactly match the land mass boundaries. 
The result of this missmatch is that at least one of the environmental variables was not available for `r format(complete_climate_cells[1], big.mark = ",")` of the raster cells with occurrences (`r scales::percent(complete_climate_cells[1]/sum(complete_climate_cells), accuracy = 0.1)` of the total). 
As expected a vast majority of these cells were close to the shore. 
To address this issue, calculated average value of environmental variables within a `r config$climate_buffer/1000`km buffer of the center of the cell where the variable is missing, and used it to approximate the value of the variable in that cell. 
In this fashion we were able to fill environmental variables for `r scales::percent(percent_filled_round_one, accuracy = 0.1)` of the cells were they were missing. 
To fill the remaining `r format(complete_filled_cells_1[1], big.mark= ",")` cells, we repeated the aforementioned procedure but using a `r config$climate_buffer*2/1000`km buffer.

Our main objective is to investigate wether the suitability of the environment was to obtain the suitability of the environment at the locations for which we have community data. 
This involves two steps. 
First, we determine the species' grinellian niche, the range of environmental conditions that are suitable for a positive growth rate. 
We assume that the suitability of a particular set of environmental conditions (a point in the environmental space) can be inferred from the density of occurrences under these conditions. 
This is, areas of the niche that have a higher density of occurrences have higher suitability and can be interpreted as the core of the species distribution in environental space. 
Contrastingly, areas with a lower density are comparatively less suitable and represent the fringes of the species distribution. 
The second step is to determine the location of our plant-pollinator communities (where the species is present) in the environmental space and calculate its corresponding environmental suitability.

To determine the environmental niche of each species we use a principal component analysis of the 37 bioclimatic variables associated with the species occurrences. 
Specifically we use the `dudi.pca` function from the R package `ade4` `r as.character(packageVersion('ade4'))` [@dray_ade4_2007] and center and scale all bioclimatic variables to have a mean of 0 and a unit standard deviation. 
We use the first two components to define the environmental space. 
We then  The use of a kernel smoother makes the process of moving from geographical space to multivariate environmental space independent of both sam- pling effort and arbitrary choice of resolution in environmental space

We removed from further analyis occurrences located within the `r complete_filled_cells_2[1]` cells for which we were unable to fill environmental variables.
